---
title: "Review"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")

# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,nycflights13)

#Consider the data in the nycflights13 package.

#Problem 2: Is there a relationship between the age of a plane and its delays?

# Load libraries
library(tidyverse)
library(nycflights13)

# Putting both datasets together by a common indentifyer, in this case, tailnum.
merged_data <- inner_join(flights, planes, by = "tailnum")

# Getting Planes' Ages
merged_data <- mutate(merged_data, Age = 2013 - year.y)

#Seeing the relationship with a Scatterplot (with dep_delay and arr_delay)
ggplot(merged_data, aes(x = Age, y = dep_delay)) +
  geom_point(alpha = 0.3) +
  labs(title = "Departure Delay vs Plane Age",
       x = "Plane Age (years)",
       y = "Departure Delay (minutes)") 

ggplot(merged_data, aes(x = Age, y = arr_delay)) +
  geom_point(alpha = 0.3) +
  labs(title = "Arrival Delay vs Plane Age",
       x = "Plane Age (years)",
       y = "Arrival Delay (minutes)") 

#Answer: Both scatterplots have a right-skewed distribution, meaning that the newer the planes the more delays they present, both for arrivals and departures. This doesn't necessarily mean that older planes used to be better, but it could be related to the fact that newer planes probably fly more and therefore there's more place for delays to happen.

#Problem 3: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

#1st: Create a "Date" variable based on year, month and days, and then group the flights based on this date.

flights$date <- lubridate::make_date(flights$year, flights$month, flights$day)
daily_groups <- group_by(flights, date)

#2nd: Now let's count how many flights happened each day.

daily_counts <- summarise(daily_groups, num_flights = n())

#3rd: the next step would be to get the median value for each day.

daily_median <- summarise(daily_groups,num_flights = n(), median_dep_delay = median(dep_delay, na.rm = TRUE))

#4th: Now we can get the top 10 values.

top10_days <- slice_head(arrange(daily_median, desc(median_dep_delay)), n = 10)

#5th: Now let's see the results.
View(top10_days)

```
